---
title: "Getting Started with DataInsightR"
author: "Omar Hamad"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with DataInsightR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Introduction

DataInsightR is an R package designed to make exploratory data analysis easier and more automated. This vignette shows you how to use the main functions.

## Installation
```{r eval=FALSE}
# Install from GitHub
devtools::install_github("ohamad03/DataInsightR")
```

## Loading the Package
```{r eval=FALSE}
library(DataInsightR)
```

## Basic Usage

### 1. Smart Data Summary

The `smart_summary()` function gives you a complete overview of your dataset.
```{r eval=FALSE}
# Using the mtcars dataset
data(mtcars)

# Get comprehensive summary
summary_result = smart_summary(mtcars)
print(summary_result)
```

This function automatically:
- Detects numeric and categorical variables
- Calculates descriptive statistics
- Reports missing values
- Provides data quality insights

### 2. Detecting Outliers

Find unusual values in your data using multiple methods:
```{r eval=FALSE}
# Detect outliers using both IQR and Z-score methods
outliers = detect_outliers(mtcars, method = "both")
print(outliers)

# Use only IQR method
outliers_iqr = detect_outliers(mtcars, method = "iqr")
```

### 3. Missing Data Analysis

Analyze patterns in missing data:
```{r eval=FALSE}
# Using airquality dataset which has missing values
data(airquality)

missing_report = missing_analyzer(airquality)
print(missing_report)
```

### 4. Correlation Analysis

Find relationships between variables:
```{r eval=FALSE}
# Analyze correlations
corr_result = correlation_matrix(mtcars, method = "pearson")
print(corr_result)

# Set minimum correlation threshold
corr_result = correlation_matrix(mtcars, min_correlation = 0.5)
```

### 5. Automatic Plotting

Let the package choose the right plot for your data:
```{r eval=FALSE}
# Single variable plot (histogram for numeric)
plot1 = auto_plot(mtcars, "mpg")
print(plot1)

# Two variable plot (scatter for numeric vs numeric)
plot2 = auto_plot(mtcars, "hp", "mpg")
print(plot2)

# Categorical vs numeric (boxplot)
plot3 = auto_plot(mtcars, "cyl", "mpg")
print(plot3)
```

### 6. Feature Engineering

Create new features automatically:
```{r eval=FALSE}
# Create polynomial and interaction features
new_data = feature_engineer(mtcars, interactions = TRUE, polynomials = TRUE)

# Check how many new features were created
cat("Original features:", attr(new_data, "original_features"), "\n")
cat("New features:", attr(new_data, "new_features"), "\n")
```

### 7. Data Quality Assessment (S4 Class)

Use the S4 class for comprehensive quality scoring:
```{r eval=FALSE}
# Assess data quality
quality = assess_data_quality(iris)
show(quality)

# Get the quality score
score = getScore(quality)
cat("Quality Score:", score, "\n")
```

### 8. Distribution Fitting

Test which statistical distributions fit your data:
```{r eval=FALSE}
# Test if data follows normal, uniform, or exponential distribution
dist_result = distribution_fit(mtcars$mpg)
print(dist_result)
```

## Complete Workflow Example

Here's a typical workflow combining multiple functions:
```{r eval=FALSE}
# Load data
data(mtcars)

# Step 1: Get overview
summary = smart_summary(mtcars)

# Step 2: Check quality
quality = assess_data_quality(mtcars)
cat("Quality score:", getScore(quality), "\n")

# Step 3: Look for outliers
outliers = detect_outliers(mtcars)

# Step 4: Analyze relationships
correlations = correlation_matrix(mtcars, min_correlation = 0.6)

# Step 5: Visualize key relationships
plot = auto_plot(mtcars, "wt", "mpg")
print(plot)
```

## Tips and Best Practices

1. Always start with `smart_summary()` to understand your data
2. Check data quality before analysis using `assess_data_quality()`
3. Use `detect_outliers()` to find unusual values that might need investigation
4. The `auto_plot()` function saves time by choosing appropriate visualizations
5. Use `feature_engineer()` before machine learning to create useful predictors

## Summary

DataInsightR provides a comprehensive toolkit for exploratory data analysis. The package implements both S3 and S4 object-oriented programming, includes defensive programming checks, and automates many common data analysis tasks.

For more information, visit the GitHub repository: https://github.com/ohamad03/DataInsightR